Proof nets for first-order additive linear logic

### 2017-12-22 ALL1.tex
lutz' comments

@@

A \defn{witness map} $\sigma\colon \varE \rightharpoonup \terms$ is a substitution map which assigns terms to existential variables, given as a (finite) partial function. Its \defn{domain} $\dom(\sigma)$ are those variables for which it is defined. The \defn{empty} witness map $\varnothing$ is the everywhere undefined partial function, and the map $\sigma[t/x]$ substitutes $t$ for $x$ and is as $\sigma$ everywhere else. We express by $\sigma(x)=\bot$ that $\sigma$ is undefined for $x$. The map $\sigma\minus x$ is undefined on $x$ and otherwise as $\sigma$.

@@

A \defn{witness linking} $\lambda_\Sigma$ is a linking $\lambda$ with a \defn{witness labelling} $\Sigma\colon\lambda\to \varE\rightharpoonup\terms$ that assigns each link $(C,D)$ a witness map. We may use and define $\lambda_\Sigma$ as a set of \defn{witness links} $\link CD$, where $(C,D)$ in $\lambda$ and $\Sigma(C,D)=\sigma$. When link is not a witness link may be emphasized by calling it \defn{clean}.

@@

\begin{definition}
The \defn{de-sequentialization} $[\pi]$ of a sequent proof $\prf\pi AB$ is the witness pre-net $\net{\lambda_\Sigma}AB$ where $\lambda_\Sigma=\deseq[\varnothing]\pi AB$ and $\deseq-AB$ is defined as follows.
%
\newlength\deseqwidth
\settowidth\deseqwidth{$\scriptstyle{A'\,,\,B'_0\+B'_1}$}
\newcommand\XDeseq[4][\sigma]{\left[\vcenter{#2}\right]_{#1}^{\makebox[\deseqwidth][l]{$\scriptstyle{#3\,,\,#4}$}}}
\begin{align*}
	\XDeseq{\infer{\seq P{\dual P}}{}}QR &\quad=\quad \{\link QR\}
\\ \\
	\XDeseq{\infer {\seq A {B_0\+B_1}} {\deduce [\rule{0pt}{1pt}] {\seq A{B_i}} \pi }}
	  {A'}{B_0'\+B'_1} 
	& \quad=\quad 
	\Deseq{\deduce[\rule{0pt}{1pt}]{\seq A{B_i}}\pi}{A'}{B_i'}
\\ \\
	\XDeseq{\infer{\seq A{B\*C}}{
	  \deduce[\rule{0pt}{1pt}]{\seq AB}\pi 
	  && 
	  \deduce[\rule{0pt}{1pt}]{\seq AC}\Sigma}}{A'}{B'\*C'}
	& \quad=\quad 
	  \Deseq{\deduce[\rule{0pt}{1pt}]{\seq AB}\pi}{A'}{B'}
	  \cup
	  \Deseq{\deduce[\rule{0pt}{1pt}]{\seq AC}\Sigma}{A'}{C'}
\\ \\
	\XDeseq{\infer{\seq A{\exists x.B}}{\deduce[\rule{0pt}{1pt}]{\seq A{B[t/x]}}\pi}}
	  {A'}{\exists x.B'}
	& \quad=\quad 
	 \Deseq[{\sigma[t/x]}]{\deduce[\rule{0pt}{1pt}]{\seq A{B[t/x]}}\pi}
	  {A'}{B'}
\\ \\
	\XDeseq{\infer[\!\!\scriptstyle{a\,\notin\,\textsc{fv}(A)}]
	  {\seq A{\forall a.B}}{\deduce[\rule{0pt}{1pt}]{\seq AB}\pi}}{A'}{\forall a.B'}
	& \quad=\quad 
	  \Deseq{\deduce[\rule{0pt}{1pt}]{\seq AB}\pi}{A'}{B'}
\end{align*}
\end{definition}

@@

\begin{equation}
	\net{\{\link AB\}}A{\forall a.B}
 ~\scoal~
	\net{\link A{\forall a.B}}A{\forall a.B}
\tag{$\srr!$}
\end{equation}

@@

To be explicit, to generate coalescence as a rewrite relation means the following: a pre-net coalesces to another, $\net{\lambda_\Sigma} AB\scoal \net{{\lambda'_{\Sigma'}}}AB$ if and only if there is a rewrite rule
\[
	\net{\kappa_\Xi} CD\scoal\net{\kappa'_{\Xi'}} CD
\] 
such that 
\[
	C\leq A \qquad
	D\leq B \qquad
	\kappa_\Xi\subseteq\lambda_\Sigma \qquad
	\kappa'_{\Xi'}\subseteq\lambda'_{\Sigma'} \quad
	\text{and} \quad
	\lambda_\Sigma-\kappa_\Xi=\lambda'_{\Sigma'}-\kappa'_{\Xi'}~.
\]

\begin{definition}
An \all1 \defn{witnessing proof net} or \defn{witness net} is a witness pre-net $\net{\lambda_\Sigma}AB$, with $\lambda_\Sigma$ an axiom linking, that strict-coalesces. It \defn{sequentializes} to a proof $\pi$ if its initial labelling $\net{\lambda_\Sigma^\star}AB$ reduces in $(\scoal)$ to $\net{\{\link[\varnothing]AB^\pi\}}AB$.
\end{definition}

@@

\begin{lemma}
Every pre-net on a coalescence path $\net{\lambda_\Sigma}AB \scoal^* \net{\{\link[\varnothing]AB^\pi\}}AB$ from a witness net satisfies the \defn{exact coverage} property: for any link $\link CD$, the domain of $\sigma$ is exactly the variables existentially quantified in $\seq AB$ thad bind in $C$ and $D$,
\[
	\dom\sigma = \{\,x \mid C < \exists x.X \leq A\,\}
			\cup \{\,y \mid D < \exists y.Y \leq B\,\}~.
\]
\end{lemma}

@@

\begin{proof}
By induction on the coalescence path, going forward from $\net{\lambda_\Sigma}AB$ for the first property, and working backward from $\net{\{\link[\varnothing]AB^\pi\}}AB$ for the second.
\end{proof}

@@

\begin{proof}
Let $L=\net{\lambda_\Sigma}AB$ and $R=\net{\kappa_\Xi}AB$. We will consider the four coalescence steps as shown below. For slice-correctness, we will demonstrate that the left-hand side and right-hand side of each rule belong to the same slice of $\seq AB$, or in the case of $\srr?$, naturally corresponding slices.
%
\begin{align}
  	\net{\{\link C{D_i}\}}C{D_0{\+}D_1} 
 &~\scoal~
	\net{\{\link C{D_0{\+}D_1} \}}C{D_0{\+}D_1}
\tag{$\srr+i$}
\\	\net{\{\link C{D_0},\link C{D_1}\}}C{D_0{\*}D_1}
 &~\scoal~ 
	\net{\{\link C{D_0{\*}D_1}\}}C{D_0{\*}D_1}
\tag{$\srr*$}
\\	\net{\{\link CD\}}C{\exists x.D}
 &~\scoal~ 
	\net{\{\link[\sigma\minus x]C{\exists x.D}\}}C{\exists x.D}
\tag{$\srr?$}
\\	\net{\{\link CD\}}C{\forall a.D}
 &~\scoal~
	\net{\{\link C{\forall a.D}\}}C{\forall a.D}
\tag{$\srr!$}
\end{align}
%
\begin{description}
	\item[$(\srr+i)$]
A slice $S_B$ of $B$ and $\varnothing$ containing either of $(D_0,\tau)$, $(D_1,\tau)$, and $(D_0\+D_1,\tau)$ must also contain the other two. A slice $S$ of $\seq AB$ then contains all three of $\link C{D_0}$, $\link C{D_1}$, and  $\link C{D_0{\+}D_1}$, or none. It follows that $S\cap \lambda_\Sigma$ is a singleton if and only if $S\cap\kappa_\Xi$ is. Since other slices are unaffected, $L$ is slice-correct if and only if $R$ is. For dependency-correctness, the column of $\link C{D_i}$ is that of $\link C{D_0{\+}D_1}$ plus $(D_i,\sigma|\ex B{D_i})$ itself, which is minimal in the order $\dep$.

>>>>

% ============================== FOR COMMENTS

\newcommand{\lutz}[1]{{\color{blue}     \noindent[\![\![{\bf Lutz: }#1]\!]\!]}}
\newcommand{\todo}[1]{{\color{red}     \noindent[\![\![{\bf TODO: }#1]\!]\!]}}

@@

A \defn{witness map} $\sigma\colon \varE \rightharpoonup \terms$ is a substitution map which assigns terms to existential variables, given as a (finite) partial function. Its \defn{domain} $\dom\sigma$ are those variables for which it is defined. The \defn{empty} witness map $\varnothing$ is the everywhere undefined partial function, and the map $\sigma[t/x]$ substitutes $t$ for $x$ and is as $\sigma$ everywhere else. We express by $\sigma(x)=\bot$ that $\sigma$ is undefined for $x$. The map $\sigma\minus x$ is undefined on $x$ and otherwise as $\sigma$.

@@

A \defn{witness linking} $\lambda_\Sigma$ is a linking $\lambda$ with a \defn{witness labelling} $\Sigma\colon\lambda\to \varE\rightharpoonup\terms$ that assigns each link $(C,D)$ a witness map. We may use and define $\lambda_\Sigma$ as a set of \defn{witness links} $\link CD$, where $(C,D)$ in $\lambda$ and $\Sigma(C,D)=\sigma$. When a link is not a witness link, this may be emphasized by calling it \defn{clean}.

@@

\begin{definition}
The \defn{de-sequentialization} $[\pi]$ of a sequent proof $\prf\pi AB$ is the witness pre-net $\net{\lambda_\Sigma}AB$ where $\lambda_\Sigma=\deseq[\varnothing]\pi AB$ and $\deseq-AB$ is defined as follows.
%
\newlength\deseqwidth
\settowidth\deseqwidth{$\scriptstyle{A'\,,\,B'_0\+B'_1}$}
\newcommand\XDeseq[4][\sigma]{\left[\vcenter{#2}\right]_{#1}^{\makebox[\deseqwidth][l]{$\scriptstyle{#3\,,\,#4}$}}}
\begin{align*}
	\XDeseq{\infer{\seq P{\dual P}}{}}QR &\quad=\quad \{\link QR\}
\\ \\
	\XDeseq{\infer {\seq A {B_0\+B_1}} {\deduce [\rule{0pt}{1pt}] {\seq A{B_i}} \pi }}
	  {A'}{B_0'\+B'_1} 
	& \quad=\quad 
	\Deseq{\deduce[\rule{0pt}{1pt}]{\seq A{B_i}}\pi}{A'}{B_i'}
\\ \\
	\XDeseq{\infer{\seq A{B\*C}}{
	  \deduce[\rule{0pt}{1pt}]{\seq AB}\pi 
	  && 
	  \deduce[\rule{0pt}{1pt}]{\seq AC}\Sigma}}{A'}{B'\*C'}
	& \quad=\quad 
	  \Deseq{\deduce[\rule{0pt}{1pt}]{\seq AB}\pi}{A'}{B'}
	  \cup
	  \Deseq{\deduce[\rule{0pt}{1pt}]{\seq AC}\Sigma}{A'}{C'}
\\ \\
	\XDeseq{\infer{\seq A{\exists x.B}}{\deduce[\rule{0pt}{1pt}]{\seq A{B[t/x]}}\pi}}
	  {A'}{\exists x.B'}
	& \quad=\quad 
	 \Deseq[{\sigma[t/x]}]{\deduce[\rule{0pt}{1pt}]{\seq A{B[t/x]}}\pi}
	  {A'}{B'}
\\ \\
	\XDeseq{\infer[\!\!\scriptstyle{a\,\notin\,\textsc{fv}(A)}]
	  {\seq A{\forall a.B}}{\deduce[\rule{0pt}{1pt}]{\seq AB}\pi}}{A'}{\forall a.B'}
	& \quad=\quad 
	  \Deseq{\deduce[\rule{0pt}{1pt}]{\seq AB}\pi}{A'}{B'}
\end{align*}
\lutz{why the $\Sigma$ in the third equation?}

\lutz{I don't see where the primes come in, i.e, in no equation there is a relation between $A$ and $A'$, etc.}
\end{definition}

@@

\begin{equation}
	\net{\{\link AB\}}A{\forall a.B}
 ~\scoal~
	\net{\{\link A{\forall a.B}\}}A{\forall a.B}
\tag{$\srr!$}
\end{equation}

@@

To be explicit, to generate coalescence as a rewrite relation means the following: a pre-net coalesces to another, $\net{\lambda_\Sigma} AB\scoal \net{{\lambda'_{\Sigma'}}}AB$ if and only if there is a rewrite rule
\[
	\net{\kappa_\Xi} CD\scoal\net{\kappa'_{\Xi'}} CD
\] 
such that 
\[
	C\leq A \qquad
	D\leq B \qquad
	\kappa_\Xi\subseteq\lambda_\Sigma \qquad
	\kappa'_{\Xi'}\subseteq\lambda'_{\Sigma'} \quad
	\text{and} \quad
	\lambda_\Sigma-\kappa_\Xi=\lambda'_{\Sigma'}-\kappa'_{\Xi'}~.
\]
\lutz{I think a $\Xi$ in the index is just horrible.}

\begin{definition}
An \all1 \defn{witnessing proof net} or \defn{witness net} is a witness pre-net $\net{\lambda_\Sigma}AB$, with $\lambda_\Sigma$ an axiom linking, that strict-coalesces. It \defn{sequentializes} to a proof $\pi$ if its initial labelling $\net{\lambda_\Sigma^\star}AB$ reduces in $(\scoal)$ to $\net{\{\link[\varnothing]AB^\pi\}}AB$.
\lutz{shouldn't we say somewhere that the initial linking is just on atoms?}
\end{definition}

@@

\begin{lemma}
Every pre-net on a coalescence path $\net{\lambda_\Sigma}AB \scoal^* \net{\{\link[\varnothing]AB^\pi\}}AB$ from a witness net satisfies the \defn{exact coverage} property: for any link $\link CD$, the domain of $\sigma$ is exactly the variables existentially quantified in $\seq AB$ thad bind \lutz{???} in $C$ and $D$,
\[
	\dom\sigma = \{\,x \mid C < \exists x.X \leq A\,\}
			\cup \{\,y \mid D < \exists y.Y \leq B\,\}~.
\]
\end{lemma}

@@

\begin{proof}
By induction on the coalescence path, going forward from $\net{\lambda_\Sigma}AB$ for the first property, and working backward from $\net{\{\link[\varnothing]AB^\pi\}}AB$ for the second.
\end{proof}
\lutz{I am missing something here. In the lemma there is only one statement. So, what is the first, and what the second?}

@@

\begin{proof}
Let $L=\net{\lambda_\Sigma}AB$ and $R=\net{\kappa_\Xi}AB$. We will consider the four coalescence steps as shown below. For slice-correctness, we will demonstrate that the left-hand side and right-hand side of each rule belong to the same slice of $\seq AB$, or in the case of $\srr?$, naturally corresponding slices.
%
\begin{align}
  	\net{\{\link C{D_i}\}}C{D_0{\+}D_1} 
 &~\scoal~
	\net{\{\link C{D_0{\+}D_1} \}}C{D_0{\+}D_1}
\tag{$\srr+i$}
\\	\net{\{\link C{D_0},\link C{D_1}\}}C{D_0{\*}D_1}
 &~\scoal~ 
	\net{\{\link C{D_0{\*}D_1}\}}C{D_0{\*}D_1}
\tag{$\srr*$}
\\	\net{\{\link CD\}}C{\exists x.D}
 &~\scoal~ 
	\net{\{\link[\sigma\minus x]C{\exists x.D}\}}C{\exists x.D}
\tag{$\srr?$}
\\	\net{\{\link CD\}}C{\forall a.D}
 &~\scoal~
	\net{\{\link C{\forall a.D}\}}C{\forall a.D}
\tag{$\srr!$}
\end{align}
%
\begin{description}
	\item[$(\srr+i)$]
A slice $S_B$ of $B$ and $\varnothing$ containing either of $(D_0,\tau)$, $(D_1,\tau)$, and $(D_0\+D_1,\tau)$ must also contain the other two. \lutz{something is confusing here. It seems you are using different notions of slice. Above it says that correctness means that slices are singletons, and here you imply that slices contain more than one element.} A slice $S$ of $\seq AB$ then contains all three of $\link C{D_0}$, $\link C{D_1}$, and  $\link C{D_0{\+}D_1}$, or none. It follows that $S\cap \lambda_\Sigma$ is a singleton if and only if $S\cap\kappa_\Xi$ is. Since other slices are unaffected, $L$ is slice-correct if and only if $R$ is. For dependency-correctness, the column of $\link C{D_i}$ is that of $\link C{D_0{\+}D_1}$ plus $(D_i,\sigma|\ex B{D_i})$ itself, which is minimal in the order $\dep$.
